Embedding Intuitionistic into Classical Logic

### 2022-07-16 CSL/lipics-v2021-EIICL.tex
typos proof of lemma 35

\begin{lemma}\label{ap1}
	Let $\chi$ be a formula and $\{z_1\dots z_n\} = Z$ contain all free variables in $\chi$, then for every structure $\mathcal M = (M, I)$ there exists $\mathcal M_\chi = (M, I_\chi$) such that $p^{I_{\chi}} = p^I$ for all function and relation symbols $p$ occurring in $\chi$ and for all variable assignments $v$ we have
	\begin{itemize}
		\item if $\models\chi$ then $\mathcal M_\chi, v\models\chi^S_T$.
		\item if $\mathcal M, v\not\models\chi$ then $\mathcal M_\chi, v\not\models\chi^H_T$.
	\end{itemize}
\end{lemma}

@@

	1. $\chi = \varphi\circ\psi$ for $\circ\in\{\wedge,\vee,\to\}$. By induction hypothesis there exist $\mathcal M_{\varphi, v} = (M, I_{\varphi, v}), \mathcal M_{\psi, v} = (M, I_{\psi, v})$ for $\varphi,\psi$ as in the lemma. Then we can define $I_{\chi, v}$ as follows:

	For every symbol $p$ that occurs in $\varphi^S$ or $\varphi^H$ but not $\psi^S, \psi^H$ have $p^{I_{\chi, v}} = p^{I_{\varphi, v}}$. For every symbol $p$ that occurs in $\psi^S$ or $\psi^H$ but not $\varphi^S, \varphi^H$ have $p^{I_{\chi, v}} = p^{I_{\psi, v}}$. Otherwise $p$ already occurs in $\varphi$ so have $p^{I_{\chi, v}} = p^I$.

	Let $\circ=\to$. Suppose $\mathcal M, v\models\chi$. Then $\mathcal M, v\not\models\varphi$ or $\mathcal M\models\psi$ and therefore $\mathcal M_\chi\not\models\varphi^S_T$ or $\mathcal M_\chi\models\psi^H_T$, i.e. $\mathcal M_\chi\models\chi^S_T$. On the other hand suppose $\mathcal M, v\not\models\chi$. Then $\mathcal M, v\models\varphi$ and $\mathcal M, v\not\models\psi$ and therefore $\mathcal M_\chi, v\models \varphi^S_T$ and $\mathcal M_\chi, v\not\models\psi^S_T$, i.e. $\mathcal M_\chi\not\models\chi^H_T$. Analogous arguments work for $\circ\in\{\wedge, \vee\}$.

	2. $\chi = \forall x\varphi$. Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that $\mathcal M, v[x_1/t_1\dots x_n/t_n, m/a]\not\models\varphi[a/x]$ for all $v$ and arbitrary otherwise. Since $T$ contains all free variables occurring in $\chi$ this is a well-defined function.

	By induction hypothesis there exists a structure $\mathcal M_{\varphi[a/x]}$ as is the Lemma. Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{T\cup a}$ or $\varphi[a/x]^H_{T\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.

	Suppose $\mathcal M, v\models\chi$. Then for all $m\in M$ $\mathcal M, v[m/a]\models\varphi[a/x]$ and therefore $\mathcal M_{\chi}, v[m/a]\models\varphi[a/x]^S_{M\cup\{a\}}$ and thus $\mathcal M_{\chi}, v\models \forall x\varphi([a/x]^S_{M\cup\{a\}}[x/a])$, i.e. $\mathcal M_\chi,v\models \chi^S$. On the other hand suppose $\mathcal M, v\not\models\chi$. Then there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\varphi[a/x]$ and therefore $\mathcal M_\chi, v[m/a]\not\models\varphi[a/x]^H_{T\cup\{a\}}$ and so by definition $\mathcal M_\chi, v\not\models\varphi[s(t_1,\dots t_n)/x]^H_T$, i.e. $\mathcal M_\chi, v\not\models(\forall x\varphi)^H_T$.

	3. $\chi = \exists x\varphi$. The argument runs dually to 2. Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that for all $v$ we have $\mathcal M, v[x_1/t_1\dots x_n/t_n, m/a]\models\varphi[a/x]$ and arbitrary otherwise. Since $T$ contains all free variables occurring in $\chi$ this is a well-defined function.

	By induction hypothesis there exists a structure $\mathcal M_{\varphi[a/x]}$ as is the Lemma. Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{T\cup a}$ or $\varphi[a/x]^H_{T\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.

	Then as in 2 from $\mathcal M, v\models \chi$ follows $\mathcal M_\chi,v\models\chi^S_T$ and from $\mathcal M, v\not\models \chi$ follows $\mathcal M_\chi,v\not\models\chi^H_T$.
\end{proof}

>>>>

\begin{lemma}\label{ap1}
Let $\chi$ be a formula and let $\{z_1\dots z_n\} = Z$ contain all free variables in $\chi$, then for every structure $\mathcal M = (M, I)$ there exists $\mathcal M_\chi = (M, I_\chi$) such that $p^{I_{\chi}} = p^I$ for all function and relation symbols $p$ occurring in $\chi$ and for all variable assignments $v$ we have
	\begin{itemize}
		\item if $M, v \models\chi$ then $\mathcal M_\chi, v\models\chi^S_T$.
		\item if $\mathcal M, v\not\models\chi$ then $\mathcal M_\chi, v\not\models\chi^H_T$.
	\end{itemize}
\end{lemma}

@@

	1. $\chi = \varphi\circ\psi$ for $\circ\in\{\wedge,\vee,\to\}$. By induction hypothesis there exist $\mathcal M_{\varphi, v} = (M, I_{\varphi, v})$ and $\mathcal M_{\psi, v} = (M, I_{\psi, v})$ for $\varphi,\psi$ as stated in the lemma. 
Then we can define $I_{\chi, v}$ as follows:

	For every symbol $p$ that occurs in $\varphi^S_Z$ or $\varphi^H_Z$ but not $\psi^S_Z, \psi^H_Z$, we set $p^{I_{\chi, v}} = p^{I_{\varphi, v}}$. For every symbol $p$ that occurs in $\psi^S_Z$ or $\psi^H_Z$ but not $\varphi^S_Z, \varphi^H_Z$, we set $p^{I_{\chi, v}} = p^{I_{\psi, v}}$. Otherwise $p$ already occurs in $\varphi$ so have $p^{I_{\chi, v}} = p^I$.

	Let $\circ=\ \to$. Suppose $\mathcal M, v\models\chi$. Then $\mathcal M, v\not\models\varphi$ or $\mathcal M\models\psi$ and therefore $\mathcal M_\chi\not\models\varphi^S_Z$ or $\mathcal M_\chi\models\psi^H_Z$, i.e. $\mathcal M_\chi\models\chi^S_Z$. On the other hand suppose $\mathcal M, v\not\models\chi$. Then $\mathcal M, v\models\varphi$ and $\mathcal M, v\not\models\psi$ and therefore $\mathcal M_\chi, v\models \varphi^S_Z$ and $\mathcal M_\chi, v\not\models\psi^S_Z$, i.e. $\mathcal M_\chi\not\models\chi^H_Z$. Analogous arguments work for $\circ\in\{\wedge, \vee\}$.

	2. $\chi = \forall x\varphi$. Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that $\mathcal M, v[x_1/z_1\dots x_n/z_n, m/a]\not\models\varphi[a/x]$ for all $v$ and arbitrary otherwise. Since $Z$ contains all free variables occurring in $\chi$ this is a well-defined function.

By induction hypothesis there exists a structure $\mathcal M_{\varphi[a/x]}$ as is the Lemma. Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{Z\cup a}$ or $\varphi[a/x]^H_{Z\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.

Suppose $\mathcal M, v\models\chi$. Then, for all $m\in M$ we have that $\mathcal M, v[m/a]\models\varphi[a/x]$ and therefore $\mathcal M_{\chi}, v[m/a]\models\varphi[a/x]^S_{Z\cup\{a\}}$ and thus $\mathcal M_{\chi}, v\models \forall x\varphi([a/x]^S_{Z\cup\{a\}}[x/a])$, i.e. $\mathcal M_\chi,v\models \chi^S_Z$. On the other hand suppose $\mathcal M, v\not\models\chi$. Then there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\varphi[a/x]$ and therefore $\mathcal M_\chi, v[m/a]\not\models\varphi[a/x]^H_{Z\cup\{a\}}$ and so by definition $\mathcal M_\chi, v\not\models\varphi[s(t_1,\dots t_n)/x]^H_Z$, i.e. $\mathcal M_\chi, v\not\models(\forall x\varphi)^H_Z$.

	3. $\chi = \exists x\varphi$. The argument runs dually to 2. Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that for all $v$ we have $\mathcal M, v[x_1/z_1\dots x_n/z_n, m/a]\models\varphi[a/x]$ and arbitrary otherwise. Since $Z$ contains all free variables occurring in $\chi$ this is a well-defined function.

	By induction hypothesis there exists a structure $\mathcal M_{\varphi[a/x]}$ as is the Lemma. Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{Z\cup a}$ or $\varphi[a/x]^H_{Z\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.

	Then as in 2 from $\mathcal M, v\models \chi$ follows $\mathcal M_\chi,v\models\chi^S_Z$ and from $\mathcal M, v\not\models \chi$ follows $\mathcal M_\chi,v\not\models\chi^H_Z$.
\end{proof}
